IMPC Mouse data - Variance in sex differences

### 2019-11-06 Prep & old/prep and old- scripts/2019_Aug30-NO_VR_IMPC_variance_and_sex_difference.Rmd
Update Nov

---
title: "IMPC Mouse data - Variance in sex differences"
author: "Susanne & Felix Zajitschek"
date: "August 2019"
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 4
    toc_float: yes
  html_notebook:
    toc: yes
  pdf_document:
    toc: yes
    toc_depth: '4'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)

@@

## Set up
### Packages

@@

### Loading functions that are necessary
#### Preparation of raw data
1) Data loading and cleaning of the csv file
This step we have already done and provide a cleaned up file which is less computing intensive. However, cvs
```{r}

# loads the raw data, setting some default types for various columns

    # subset to reasonable set of variables
    # date_of_experiment: Jeremy suggested using as an indicator of batch-level effects
    select(production_center, strain_name, strain_accession_id, biological_sample_id, pipeline_stable_id, procedure_group, procedure_name, sex, date_of_experiment, age_in_days, weight, parameter_name, data_point) %>% 
    arrange(production_center, biological_sample_id, age_in_days)
}

2) Subsetting the data to choose only one datapoint per individual per trait
```{r}
# this is a necessary step for the loop across all traits 
data_subset_parameterid_individual_by_age <- function(mydata, parameter, age_min, age_center) {
  tmp <- mydata %>%
    filter(age_in_days >= age_min,
           id == parameter) %>%
    # take results for single individual closest to age_center
    mutate(age_diff = abs(age_center - age_in_days)) %>%
    group_by(biological_sample_id) %>%
    filter(age_diff == min(age_diff)) %>%
    select(-age_diff)

#### Functions for preparing the data for meta analyses
3) "Population statistics"
```{r}
calculate_population_stats <- function(mydata, min_individuals = 5) {
  mydata %>% 
    group_by(population, strain_name, production_center, sex) %>% 
    summarise(
      trait = parameter_name[1],
      x_bar = mean(data_point),
      x_sd = sd(data_point),
      n_ind = n()
    ) %>% 
    ungroup() %>%
    filter(n_ind > min_individuals) %>% 
    # Check both sexes present & filter those missing
    group_by(population) %>% 
    mutate(
      n_sex = n_distinct(sex)
    ) %>% 
    ungroup() %>%
    filter(n_sex ==2) %>% 
    select(-n_sex) %>%
    arrange(production_center, strain_name, population, sex)
}
```

 4) Extraction of effect sizes and sample variances
```{r}
create_meta_analysis_effect_sizes <- function(mydata) {
  i <- seq(1, nrow(mydata), by = 2)
  input <- data.frame(
    n1i = mydata$n_ind[i],
    n2i = mydata$n_ind[i + 1],
    x1i = mydata$x_bar[i],
    x2i = mydata$x_bar[i + 1],
    sd1i = mydata$x_sd[i],
    sd2i = mydata$x_sd[i + 1]
  )

@@

 5) Meta Analysis

@@

@@

We here provide the cleaned dataset, which we have saved in a folder called "export", as easy starting point. However, the full dataset can be loaded and cleaned using the data cleaning function (Function 1 above), if "#" signs in the code below are removed  (created as that is much smaller than the .csv - which we can still provide for those who absolutely want to start from scratch?)
```{r}
## Load raw data - save cleaned dataset as RDS for reuse
#data_raw <- load_raw("data/dr7.0_all_control_data.csv") %>%
#    clean_raw_data()
#dir.create("export", F, F)
#saveRDS(data_raw, "export/data_clean.rds")
getwd()
data1 <- readRDS("../export/data_clean.rds") 
#data1 
```

@@

## Meta analysis, Phase 1
### Preparation
#### Define population variable & add grouping variables

@@

 group <- read.csv("../export/ParameterGrouping.csv") 
 data <- data3
 data$parameterGroup <- group$parameter[match(data$parameter_name, group$parameter_name)]

#### Assign each unique parameter_name (=trait,use trait variable) a unique number ('id')
We add a new variable, where redundant traits are combined
[note however, at this stage the dataset still contains nonsensical traits, i.e. traits that may not contain any information on variance]

@@

#### Create a matrix to store results for all traits
As the current version of this script utilizes a loop instead of tidyR code, it is here necessary to create an empty matrix, in which the returning values will be stored.

@@

### LOOP, to run meta-analysis on all traits
The loop combines the functions mentioned above and fills the data matrix with results from our meta analysis. 
Error messages indicate traits that either did not reach convergence, or that did not return meaningful results in the meta-analysis, due to absence of variance. Those traits will be removed in later steps, outlined below.

@@

### Merging datasets

@@

```{r}
procedures <- read.csv("../export/procedures.csv")

@@

```{r}
meta_clean <- meta1[ !(meta1$id %in% c(84,144,158,160,161,162,163,165,166,167,168,221,222,231)), ]
removed <-length(unique(meta_clean$parameter_name)) #218
```

## Meta-analysis, Phase 2
### Dealing with Correlated Parameters, preparation

This dataset contained a number of highly correlated traits, such as different kinds of cell counts (for example, hierarchical parametrization within immunological assays). As those data-points are not independent of each other,  we conducted a meta analyses on these correlated parameters to collapse the number of levels.

@@

meta1 <- meta_clean 
# length(unique(meta1$procedure)) #18
# length(unique(meta1$GroupingTerm)) #9 
# length(unique(meta1$parameter_group)) # 148 levels. To be used as grouping factor for meta-meta analysis / collapsing down based on things that are classified identically in "parameter_group" but have different "parameter_name" 
length(unique(meta1$parameter_name)) #218

@@

#### Count of number of parameter names (correlated sub-traits) in each parameter group (par_group_size) 
This serves to identify and separate the traits that are correlated from the full dataset that can be processed as is. If the sample size (n) for a given "parameter group" equals 1, the trait is unique and uncorrelated. All instances, where there are 2 or more traits associated with the same  parameter group (90 cases), are selected for a "mini-meta analysis", which removes the issue of correlation.

@@

### Perform meta-analyses on correlated sub-traits, using `robumeta` 
The subset of the data is prepared (nested), and in this first step the model of the meta analysis effect sizes are calculated

@@

#### Extract and save parameter estimates 
Function to collect the outcomes of the "mini" meta analysis
```{r}
count_fun <- function(mod_sub)
  return(c(mod_sub$reg_table$b.r, mod_sub$reg_table$CI.L, mod_sub$reg_table$CI.U, mod_sub$reg_table$SE))   #estimate, lower ci, upper ci, SE
```

Extraction of values created during Meta analysis using robu meta
```{r}
robusub_RR <- model_count %>% 
  transmute(parameter_group, estimatelnRR = map(model_lnRR, count_fun)) %>% 
  mutate(r = map(estimatelnRR, ~ data.frame(t(.)))) %>%
  unnest(r) %>%
  select(-estimatelnRR) %>%
  purrr::set_names(c("parameter_group","lnRR","lnRR_lower","lnRR_upper","lnRR_se"))

@@

Combine data 
```{r}
# Step1 
combinedmeta <- bind_rows(robu_all, meta_all)
#glimpse(combinedmeta)

@@

#### Clean-up and rename 
```{r}
metacombo <-metacombo[, c(1, 21:23, 2:13)] 
names(metacombo)[3] <- "procedure" 
names(metacombo)[4] <- "GroupingTerm"

@@

## Meta-analysis, Phase 3
#### Perform meta-analyses (3 for each of the 9 grouping terms: lnCVR, lnVR, lnRR) 
This is the full result dataset
```{r}

@@

### Visualisation

#### Plot FIGURE 2 [4 in ms] (First-order meta analysis results)

@@

meta.plot2.all.b <- gather(meta.plot2.all, trait, value, c(lnCVR, lnVR, lnRR))

meta.plot2.all.b$trait <- factor(meta.plot2.all.b$trait, levels =c("lnCVR","lnVR","lnRR") )

@@

meta.plot2.sig.b <- meta.plot2.sig[, c("lnCVR", "lnVR", "lnRR", "lnCVRsig", "lnVRsig", "lnRRsig", "GroupingTerm")]

@@

meta.plot2.sig.c$trait <- factor(meta.plot2.sig.c$trait, levels =c("lnCVR","lnVR","lnRR") )

@@

Prepare data for traits with effect size ratios > 10% larger in males
```{r}
meta.plot2.over10 <- meta_clean %>% select(lnCVR, lnVR, lnRR, GroupingTerm) %>% arrange(GroupingTerm)

meta.plot2.over10.b <- gather(meta.plot2.over10, trait, value, c(lnCVR, lnVR, lnRR))

meta.plot2.over10.b$trait <- factor(meta.plot2.over10.b$trait, levels =c("lnCVR","lnVR","lnRR") )

@@

Fig2 <- ggarrange(malebias_Fig2_alltraits, malebias_Fig2_sigtraits,malebias_Fig2_over10, nrow = 3, align = "v", heights = c(1.22,1,1), labels = c("A", "B", "C"))
Fig2
ggsave("Fig2.pdf", plot = Fig2, width = 6, height = 5) 
```

###  Overall results of second order meta anlaysis (Figure 4a)
#### Restructure data for plotting 
Data are restructured, and grouping terms are being re-ordered
```{r}
overall3 <- gather(overall2, parameter, value, c(lnCVR, lnVR, lnRR), factor_key= TRUE)

lnCVR.ci <- overall3 %>% filter(parameter == "lnCVR") %>% mutate(ci.low = lnCVR_lower, ci.high = lnCVR_upper)
lnVR.ci <- overall3 %>% filter(parameter == "lnVR") %>% mutate(ci.low = lnVR_lower, ci.high = lnVR_upper)
lnRR.ci <- overall3 %>% filter(parameter == "lnRR") %>% mutate(ci.low = lnRR_lower, ci.high = lnRR_upper)

overall4 <- bind_rows(lnCVR.ci, lnVR.ci, lnRR.ci) %>% select(GroupingTerm, parameter, value, ci.low, ci.high)

@@

#### Plot FIGURE 4 (Second-order meta analysis results)

Preparation: Sub-Plot  for Figure 3: all traits
```{r}

@@

########

#### Figure 4B:  Prepare data for traits with CI not overlapping 0

create column with 1= different from zero, 0= zero included in CI
Male-biased (significant) traits
```{r}
meta.male.plot3.sig <- metacombo %>% 
    	mutate(sigCVR = ifelse(lnCVR_lower > 0, 1, 0),
	sigVR = ifelse(lnVR_lower > 0, 1, 0),
	sigRR = ifelse(lnRR_lower > 0, 1, 0))

@@

#Significant subset for lnVR
metacombo_male.plot3.VR <- meta.male.plot3.sig %>%
 filter(sigVR == 1) %>%
 group_by(GroupingTerm) %>%
 nest()

metacombo_male.plot3.VR.all <- meta.male.plot3.sig %>%
 filter(sigVR == 1) %>%
 nest()

@@

plot3.male.meta.VR <- metacombo_male.plot3.VR %>% 
  mutate(model_lnVR = map(data, ~ metafor::rma.uni(yi = .x$lnVR, sei = (.x$lnVR_upper - .x$lnVR_lower)/ (2*1.96),  
				control=list(optimizer="optim", optmethod="Nelder-Mead", maxit= 1000), verbose=F)) )

@@

plot3.male.meta.VR.all <- metacombo_male.plot3.VR.all %>% 
  mutate(model_lnVR = map(data, ~ metafor::rma.uni(yi = .x$lnVR, sei = (.x$lnVR_upper - .x$lnVR_lower)/ (2*1.96),  
				control=list(optimizer="optim", optmethod="Nelder-Mead", maxit= 1000), verbose=F)) )

plot3.male.meta.VR.all <- plot3.male.meta.VR.all %>% mutate(GroupingTerm = "All")

@@

plot3.male.meta.CVR <- bind_rows(plot3.male.meta.CVR, plot3.male.meta.CVR.all)
plot3.male.meta.VR <- bind_rows(plot3.male.meta.VR, plot3.male.meta.VR.all)
plot3.male.meta.RR <- bind_rows(plot3.male.meta.RR, plot3.male.meta.RR.all)

@@

plot3.male.meta.VR.b <- as.data.frame(plot3.male.meta.VR %>% group_by(GroupingTerm) %>%  
	mutate(lnVR = map_dbl(model_lnVR, pluck(2)), lnVR_lower = map_dbl(model_lnVR, pluck(6)), 
	lnVR_upper =map_dbl(model_lnVR, pluck(7)), lnVR_se =map_dbl(model_lnVR, pluck(3))) )[, c(1,4:7)] 
plot3.male.meta.VR.b <- plot3.male.meta.VR.b[order(plot3.male.meta.VR.b$GroupingTerm),]

@@

overall.male.plot3 <- inner_join(plot3.male.meta.CVR.b, plot3.male.meta.VR.b) 
overall.male.plot3 <- inner_join(overall.male.plot3, plot3.male.meta.RR.b)

@@

Restructure MALE data for plotting 
```{r}
overall3.male.sig <- gather(overall.male.plot3, parameter, value, c(lnCVR, lnVR, lnRR), factor_key= TRUE)

lnCVR.ci <- overall3.male.sig  %>% filter(parameter == "lnCVR") %>% mutate(ci.low = lnCVR_lower, ci.high = lnCVR_upper)
lnVR.ci <- overall3.male.sig  %>% filter(parameter == "lnVR") %>% mutate(ci.low = lnVR_lower, ci.high = lnVR_upper)
lnRR.ci <- overall3.male.sig  %>% filter(parameter == "lnRR") %>% mutate(ci.low = lnRR_lower, ci.high = lnRR_upper)

overall4.male.sig <- bind_rows(lnCVR.ci, lnVR.ci, lnRR.ci) %>% select(GroupingTerm, parameter, value, ci.low, ci.high)

@@

Plot Fig3 all significant results (CI not overlapping zero) for males
```{r}
######

@@

#######################
#### Female Figure, significant traits
Female Fig3 sig

@@

```{r}
overall3.female.sig <- gather(overall.female.plot3, parameter, value, c(lnCVR, lnVR, lnRR), factor_key= TRUE)

lnCVR.ci <- overall3.female.sig  %>% filter(parameter == "lnCVR") %>% mutate(ci.low = lnCVR_lower, ci.high = lnCVR_upper)
lnVR.ci <- overall3.female.sig  %>% filter(parameter == "lnVR") %>% mutate(ci.low = lnVR_lower, ci.high = lnVR_upper)
lnRR.ci <- overall3.female.sig  %>% filter(parameter == "lnRR") %>% mutate(ci.low = lnRR_lower, ci.high = lnRR_upper)

overall4.female.sig <- bind_rows(lnCVR.ci, lnVR.ci, lnRR.ci) %>% select(GroupingTerm, parameter, value, ci.low, ci.high)

@@

#Metameta_Fig3_female.sig
```
#######################

@@

#### Male Fig 3 > 10% (male biased traits)
```{r}
meta.male.plot3.perc <- metacombo %>% 
    	mutate(percCVR = ifelse(lnCVR > log (11/10), 1, 0),			
	percVR = ifelse(lnVR > log (11/10), 1, 0),
	percRR = ifelse(lnRR > log (11/10), 1, 0))

@@

overall.male.plot3.perc <- full_join(plot3.male.meta.CVR.perc.b, plot3.male.meta.VR.perc.b) 
overall.male.plot3.perc <- full_join(overall.male.plot3.perc, plot3.male.meta.RR.perc.b)

@@

Restructure data for plotting : Male biased, 10% difference
```{r}
overall3.perc <- gather(overall.male.plot3.perc, parameter, value, c(lnCVR, lnVR, lnRR), factor_key= TRUE)

lnCVR.ci <- overall3.perc %>% filter(parameter == "lnCVR") %>% mutate(ci.low = lnCVR_lower, ci.high = lnCVR_upper)
lnVR.ci <- overall3.perc  %>% filter(parameter == "lnVR") %>% mutate(ci.low = lnVR_lower, ci.high = lnVR_upper)
lnRR.ci <- overall3.perc  %>% filter(parameter == "lnRR") %>% mutate(ci.low = lnRR_lower, ci.high = lnRR_upper)

overall4.male.perc <- bind_rows(lnCVR.ci, lnVR.ci, lnRR.ci) %>% select(GroupingTerm, parameter, value, ci.low, ci.high)

@@

Plot Fig3 all >10% difference (male bias)
```{r}

@@

##################

Female Fig 3 >10%
```{r}

@@

overall.plot3.perc <- full_join(plot3.meta.CVR.perc.b, plot3.meta.VR.perc.b) 
overall.plot3.perc <- full_join(overall.plot3.perc, plot3.meta.RR.perc.b)

@@

#### Restructure data for plotting
Female bias, 10 percent difference
```{r}
overall3.perc <- gather(overall.plot3.perc, parameter, value, c(lnCVR, lnVR, lnRR), factor_key= TRUE)

lnCVR.ci <- overall3.perc %>% filter(parameter == "lnCVR") %>% mutate(ci.low = lnCVR_lower, ci.high = lnCVR_upper)
lnVR.ci <- overall3.perc  %>% filter(parameter == "lnVR") %>% mutate(ci.low = lnVR_lower, ci.high = lnVR_upper)
lnRR.ci <- overall3.perc  %>% filter(parameter == "lnRR") %>% mutate(ci.low = lnRR_lower, ci.high = lnRR_upper)

overall4.perc <- bind_rows(lnCVR.ci, lnVR.ci, lnRR.ci) %>% select(GroupingTerm, parameter, value, ci.low, ci.high)

@@

#### LOOP
Parameters to extract from metafor (sigma2's, s.nlevels)

@@

#### Last step: meta-meta-analysis of heterogeneity
```{r}
## *Perform meta-meta-analysis (3 for each of the 9 grouping terms: var.CVR, var.VR, var.RR)

@@

>>>>

---
title: "IMPC Mouse data - Variance in sex differences"
author: "Susanne Zajitschek,  Felix Zajitschek, Russell Bonduriansky,Robert  Brooks, Will Cornwell, Daniel Falster, Malgortaza Lagisz, Jeremy Mason, Daniel Noble, Alistair Senior & Shinichi Nakagawa"
date: "August 2019"
output:
  html_document:
    code_download: true
    code_folding: hide
    depth: 4
    number_sections: no
    theme:  flatly
    toc: yes
    toc_depth: 4
    toc_float: yes
  html_notebook:
    toc: yes
  pdf_document:
    toc: yes
    toc_depth: '4'
subtitle: Electronic Supplementary Material
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE,
                      cache = TRUE,
                      tidy = TRUE)

@@

## Set up
#### Packages

@@

### Loading functions 
#### Functions for preparing the data for meta analyses
1) "Population statistics"
```{r}
calculate_population_stats <- function(mydata, min_individuals = 5) {
  mydata %>% 
    group_by(population, strain_name, production_center, sex) %>% 
    summarise(
      trait = parameter_name[1],
      x_bar = mean(data_point),
      x_sd = sd(data_point),
      n_ind = n()
    ) %>% 
    ungroup() %>%
    filter(n_ind > min_individuals) %>% 
    # Check both sexes present & filter those missing
    group_by(population) %>% 
    mutate(
      n_sex = n_distinct(sex)
    ) %>% 
    ungroup() %>%
    filter(n_sex ==2) %>% 
    select(-n_sex) %>%
    arrange(production_center, strain_name, population, sex)
}
```

2) Extraction of effect sizes and sample variances
```{r}
create_meta_analysis_effect_sizes <- function(mydata) {
  i <- seq(1, nrow(mydata), by = 2)
  input <- data.frame(
    n1i = mydata$n_ind[i],
    n2i = mydata$n_ind[i + 1],
    x1i = mydata$x_bar[i],
    x2i = mydata$x_bar[i + 1],
    sd1i = mydata$x_sd[i],
    sd2i = mydata$x_sd[i + 1]
  )

@@

3) Meta Analysis

@@

## The dataset

@@

We here provide the cleaned dataset, which we have saved in a folder called "export", as easy starting point. However, the full dataset can be loaded and cleaned using the data cleaning function (Function 1 above), if "#" signs in the code below are removed  (created as that is much smaller than the .csv - which we can still provide for those who absolutely want to start from scratch?)

#### Preparation of raw data
1) Data loading and cleaning of the csv file
This step we have already done and provide a cleaned up file which is less computing intensive. However, the cvs is provided in case this is pereferred to be attempted, following the steps below:

```{r Raw Data, eval=FALSE, include=TRUE}
# loads the raw data, setting some default types for various columns

    # subset to reasonable set of variables
    # date_of_experiment: Jeremy suggested using as an indicator of batch-level effects
    select(production_center, strain_name, strain_accession_id, biological_sample_id, pipeline_stable_id, procedure_group, procedure_name, sex, date_of_experiment, age_in_days, weight, parameter_name, data_point) %>% 
    arrange(production_center, biological_sample_id, age_in_days)
}
```
2) Subsetting the data to choose only one data point per individual per trait
```{r}
#2) Subsetting the data to choose only one data point per individual per trait

# this is a necessary step for the loop across all traits 
data_subset_parameterid_individual_by_age <- function(mydata, parameter, age_min, age_center) {
  tmp <- mydata %>%
    filter(age_in_days >= age_min,
           id == parameter) %>%
    # take results for single individual closest to age_center
    mutate(age_diff = abs(age_center - age_in_days)) %>%
    group_by(biological_sample_id) %>%
    filter(age_diff == min(age_diff)) %>%
    select(-age_diff)

### Load data (raw / pre-clean)

```{r}
## Load raw data - save cleaned dataset as RDS for reuse
#data_raw <- load_raw("data/dr7.0_all_control_data.csv") %>%
#    clean_raw_data()
#dir.create("export", F, F)
#saveRDS(data_raw, "export/data_clean.rds")
getwd()
data1 <- readRDS("export/data_clean.rds") 
#data1 
```

@@

## Meta analysis, Phase 1: across all traits
### Preparation
####Define population variable 
& add grouping variables

@@

 group <- read.csv("export/ParameterGrouping.csv") 
 data <- data3
 data$parameterGroup <- group$parameter[match(data$parameter_name, group$parameter_name)] 
```

#### Assign unique IDs (per trait)
each unique parameter_name (=trait,use trait variable) gets a unique number ('id')

We add a new variable, where redundant traits are combined
[note however, at this stage the dataset still contains nonsensical traits, i.e. traits that may not contain any information on variance]

@@

#### Create matrix 
As the current version of this script utilizes a loop instead of tidyR code, it is here necessary to create an empty matrix, in which the returning values will be stored.

@@

#### Loop running meta-analysis on all traits
The loop combines the functions mentioned above and fills the data matrix with results from our meta analysis. 
Error messages indicate traits that either did not reach convergence, or that did not return meaningful results in the meta-analysis, due to absence of variance. Those traits will be removed in later steps, outlined below.

@@

#### Merging datasets & removal of non-converged traits

@@

```{r}
procedures <- read.csv("export/procedures.csv")

@@

```{r}
meta_clean <- meta1[ !(meta1$id %in% c(84,144,158,160,161,162,163,165,166,167,168,221,222,231)), ]
removed <-length(unique(meta_clean$parameter_name)) #218

meta_problem <- results.alltraits.grouping[ (results.alltraits.grouping$id %in% c(84,144,158,160,161,162,163,165,166,167,168,221,222,231)), ]
```

## Meta-analysis, Phase 2: non-independent traits
### Dealing with Correlated Parameters, preparation

This dataset contained a number of highly correlated traits, such as different kinds of cell counts (for example, hierarchical parameterization within immunological assays). As those data-points are not independent of each other,  we conducted a meta analyses on these correlated parameters to collapse the number of levels.

@@

meta1 <- meta_clean 
# length(unique(meta1$procedure)) #18
# length(unique(meta1$GroupingTerm)) #9 
# length(unique(meta1$parameter_group)) # 148 levels. To be used as grouping factor for meta-meta analysis / collapsing down based on things that are classified identically in "parameter_group" but have different "parameter_name" 
#length(unique(meta1$parameter_name)) #218

@@

#### Count of number of parameter names (correlated sub-traits) in each parameter group (par_group_size)

This serves to identify and separate the traits that are correlated from the full dataset that can be processed as is. If the sample size (n) for a given "parameter group" equals 1, the trait is unique and uncorrelated. All instances, where there are 2 or more traits associated with the same  parameter group (90 cases), are selected for a "mini-meta analysis", which removes the issue of correlation.

@@

#### Meta-analyses on correlated (sub-)traits, using robumeta` 
The subset of the data is prepared (nested), and in this first step the model of the meta analysis effect sizes are calculated

@@

Extract and save parameter estimates:
Function to collect the outcomes of the "mini" meta analysis
```{r}
count_fun <- function(mod_sub)
  return(c(mod_sub$reg_table$b.r, mod_sub$reg_table$CI.L, mod_sub$reg_table$CI.U, mod_sub$reg_table$SE))   #estimate, lower ci, upper ci, SE
```

Extraction of values created during Meta analysis using robu meta:
```{r}
robusub_RR <- model_count %>% 
  transmute(parameter_group, estimatelnRR = map(model_lnRR, count_fun)) %>% 
  mutate(r = map(estimatelnRR, ~ data.frame(t(.)))) %>%
  unnest(r) %>%
  select(-estimatelnRR) %>%
  purrr::set_names(c("parameter_group","lnRR","lnRR_lower","lnRR_upper","lnRR_se"))

@@

#### Combine data 
```{r}
# Step1 
combinedmeta <- bind_rows(robu_all, meta_all)
#glimpse(combinedmeta)

@@

 Clean-up and rename 
```{r}
metacombo <-metacombo[, c(1, 21:23, 2:13)] 
names(metacombo)[3] <- "procedure" 
names(metacombo)[4] <- "GroupingTerm"

@@

## Meta-analysis, Phase 3: Second-order meta analysis for functional groups
#### Perform meta-analyses (3 for each of the 9 grouping terms: lnCVR, lnVR, lnRR) 
This is the full result dataset
```{r}

@@

## Visualisation

#### Plot:  FIGURE 2 [Figure 4 in ms] (First-order meta analysis results)

@@

meta.plot2.all.b <- gather(meta.plot2.all, trait, value, c(lnCVR,  lnRR)) #lnVR,

meta.plot2.all.b$trait <- factor(meta.plot2.all.b$trait, levels =c("lnCVR","lnRR") )#"lnVR",

@@

meta.plot2.sig.b <- meta.plot2.sig[, c("lnCVR",  "lnRR", "lnCVRsig", "lnVRsig", "lnRRsig", "GroupingTerm")]   #"lnVR",

@@

meta.plot2.sig.c$trait <- factor(meta.plot2.sig.c$trait, levels =c("lnCVR","lnRR") ) #"lnVR",

@@

Prepare data for traits with effect size ratios > 10% larger in males
```{r}
meta.plot2.over10 <- meta_clean %>% select(lnCVR,  lnRR, GroupingTerm) %>% arrange(GroupingTerm) #lnVR,

meta.plot2.over10.b <- gather(meta.plot2.over10, trait, value, c(lnCVR,  lnRR)) #lnVR,

meta.plot2.over10.b$trait <- factor(meta.plot2.over10.b$trait, levels =c("lnCVR","lnRR") ) #"lnVR",

@@

Fig2 <- ggarrange(malebias_Fig2_alltraits, malebias_Fig2_sigtraits,malebias_Fig2_over10, nrow = 3, align = "v", heights = c(1.22,1,1), labels = c("A", "B", "C"))
Fig2
ggsave("CountDataCVR_RR_only.pdf", plot = Fig2, width = 6, height = 5) 
```

###  Overall results of second order meta analysis (Figure 4a)
#### Restructure data for plotting 
Data are restructured, and grouping terms are being re-ordered
```{r}
overall3 <- gather(overall2, parameter, value, c(lnCVR,  lnRR), factor_key= TRUE) #lnVR,

lnCVR.ci <- overall3 %>% filter(parameter == "lnCVR") %>% mutate(ci.low = lnCVR_lower, ci.high = lnCVR_upper)
#lnVR.ci <- overall3 %>% filter(parameter == "lnVR") %>% mutate(ci.low = lnVR_lower, ci.high = lnVR_upper)
lnRR.ci <- overall3 %>% filter(parameter == "lnRR") %>% mutate(ci.low = lnRR_lower, ci.high = lnRR_upper)

overall4 <- bind_rows(lnCVR.ci,  lnRR.ci) %>% select(GroupingTerm, parameter, value, ci.low, ci.high) #lnVR.ci,

@@

#### Plot FIGURE 4 (Second-order meta analysis results)
Preparation: Sub-Plot  for Figure 3: all traits
```{r}

@@

#### Figure 4B- traits with CI not overlapping 0 
Prepare data 
create column with 1= different from zero, 0= zero included in CI
Male-biased (significant) traits
```{r}
meta.male.plot3.sig <- metacombo %>% 
    	mutate(sigCVR = ifelse(lnCVR_lower > 0, 1, 0),
	sigVR = ifelse(lnVR_lower > 0, 1, 0),
	sigRR = ifelse(lnRR_lower > 0, 1, 0))

@@

#Significant subset for lnVR
#metacombo_male.plot3.VR <- meta.male.plot3.sig %>%
# filter(sigVR == 1) %>%
# group_by(GroupingTerm) %>%
# nest()

#metacombo_male.plot3.VR.all <- meta.male.plot3.sig %>%
 #filter(sigVR == 1) %>%
 #nest()

@@

#plot3.male.meta.VR <- metacombo_male.plot3.VR %>% 
 # mutate(model_lnVR = map(data, ~ metafor::rma.uni(yi = .x$lnVR, sei = (.x$lnVR_upper - .x$lnVR_lower)/ (2*1.96),  
#				control=list(optimizer="optim", optmethod="Nelder-Mead", maxit= 1000), verbose=F)) )

@@

#plot3.male.meta.VR.all <- metacombo_male.plot3.VR.all %>% 
#  mutate(model_lnVR = map(data, ~ metafor::rma.uni(yi = .x$lnVR, sei = (.x$lnVR_upper - .x$lnVR_lower)/ (2*1.96),  
#				control=list(optimizer="optim", optmethod="Nelder-Mead", maxit= 1000), verbose=F)) )

#plot3.male.meta.VR.all <- plot3.male.meta.VR.all %>% mutate(GroupingTerm = "All")

@@

plot3.male.meta.CVR <- bind_rows(plot3.male.meta.CVR, plot3.male.meta.CVR.all)
#plot3.male.meta.VR <- bind_rows(plot3.male.meta.VR, plot3.male.meta.VR.all)
plot3.male.meta.RR <- bind_rows(plot3.male.meta.RR, plot3.male.meta.RR.all)

@@

#plot3.male.meta.VR.b <- as.data.frame(plot3.male.meta.VR %>% group_by(GroupingTerm) %>%  
#	mutate(lnVR = map_dbl(model_lnVR, pluck(2)), lnVR_lower = map_dbl(model_lnVR, pluck(6)), 
#	lnVR_upper =map_dbl(model_lnVR, pluck(7)), lnVR_se =map_dbl(model_lnVR, pluck(3))) )[, c(1,4:7)] 
#plot3.male.meta.VR.b <- plot3.male.meta.VR.b[order(plot3.male.meta.VR.b$GroupingTerm),]

@@

#overall.male.plot3 <- full_join(plot3.male.meta.CVR.b, plot3.male.meta.VR.b) 
overall.male.plot3 <- full_join(plot3.male.meta.CVR.b, plot3.male.meta.RR.b)

@@

Restructure MALE data for plotting 
```{r}
overall3.male.sig <- gather(overall.male.plot3, parameter, value, c(lnCVR,  lnRR), factor_key= TRUE) #lnVR,

lnCVR.ci <- overall3.male.sig  %>% filter(parameter == "lnCVR") %>% mutate(ci.low = lnCVR_lower, ci.high = lnCVR_upper)
#lnVR.ci <- overall3.male.sig  %>% filter(parameter == "lnVR") %>% mutate(ci.low = lnVR_lower, ci.high = lnVR_upper)
lnRR.ci <- overall3.male.sig  %>% filter(parameter == "lnRR") %>% mutate(ci.low = lnRR_lower, ci.high = lnRR_upper)

overall4.male.sig <- bind_rows(lnCVR.ci,  lnRR.ci) %>% select(GroupingTerm, parameter, value, ci.low, ci.high) #lnVR.ci,

@@

Plot Fig3 all significant results (CI not overlapping zero) for males
```{r}

@@

#### Female Figure, significant traits
Female Fig3 sig

@@

```{r}
overall3.female.sig <- gather(overall.female.plot3, parameter, value, c(lnCVR,  lnRR), factor_key= TRUE) #lnVR,

lnCVR.ci <- overall3.female.sig  %>% filter(parameter == "lnCVR") %>% mutate(ci.low = lnCVR_lower, ci.high = lnCVR_upper)
#lnVR.ci <- overall3.female.sig  %>% filter(parameter == "lnVR") %>% mutate(ci.low = lnVR_lower, ci.high = lnVR_upper)
lnRR.ci <- overall3.female.sig  %>% filter(parameter == "lnRR") %>% mutate(ci.low = lnRR_lower, ci.high = lnRR_upper)

overall4.female.sig <- bind_rows(lnCVR.ci,  lnRR.ci) %>% select(GroupingTerm, parameter, value, ci.low, ci.high) #lnVR.ci,

@@

#Metameta_Fig3_female.sig
```

@@

#### Male Fig 4 > 10% (male biased traits)
```{r}
meta.male.plot3.perc <- metacombo %>% 
    	mutate(percCVR = ifelse(lnCVR > log (11/10), 1, 0),			
	percVR = ifelse(lnVR > log (11/10), 1, 0),
	percRR = ifelse(lnRR > log (11/10), 1, 0))

@@

#overall.male.plot3.perc <- full_join(plot3.male.meta.CVR.perc.b, plot3.male.meta.VR.perc.b) 
overall.male.plot3.perc <- full_join(plot3.male.meta.CVR.perc.b, plot3.male.meta.RR.perc.b)

@@

Restructure data for plotting : Male biased, 10% difference
```{r}
overall3.perc <- gather(overall.male.plot3.perc, parameter, value, c(lnCVR, lnRR), factor_key= TRUE) #lnVR,

lnCVR.ci <- overall3.perc %>% filter(parameter == "lnCVR") %>% mutate(ci.low = lnCVR_lower, ci.high = lnCVR_upper)
#lnVR.ci <- overall3.perc  %>% filter(parameter == "lnVR") %>% mutate(ci.low = lnVR_lower, ci.high = lnVR_upper)
lnRR.ci <- overall3.perc  %>% filter(parameter == "lnRR") %>% mutate(ci.low = lnRR_lower, ci.high = lnRR_upper)

overall4.male.perc <- bind_rows(lnCVR.ci,  lnRR.ci) %>% select(GroupingTerm, parameter, value, ci.low, ci.high) #lnVR.ci,

@@

Plot Fig3 (4 in ms) all >10% difference (male bias)
```{r}

@@

####Female Fig 3 >10%
```{r}

@@

#overall.plot3.perc <- full_join(plot3.meta.CVR.perc.b, plot3.meta.VR.perc.b) 
overall.plot3.perc <- full_join(plot3.meta.CVR.perc.b, plot3.meta.RR.perc.b)

@@

Restructure data for plotting
Female bias, 10 percent difference
```{r}
overall3.perc <- gather(overall.plot3.perc, parameter, value, c(lnCVR, lnRR), factor_key= TRUE) #lnVR,

lnCVR.ci <- overall3.perc %>% filter(parameter == "lnCVR") %>% mutate(ci.low = lnCVR_lower, ci.high = lnCVR_upper)
#lnVR.ci <- overall3.perc  %>% filter(parameter == "lnVR") %>% mutate(ci.low = lnVR_lower, ci.high = lnVR_upper)
lnRR.ci <- overall3.perc  %>% filter(parameter == "lnRR") %>% mutate(ci.low = lnRR_lower, ci.high = lnRR_upper)

overall4.perc <- bind_rows(lnCVR.ci,  lnRR.ci) %>% select(GroupingTerm, parameter, value, ci.low, ci.high) #lnVR.ci,

@@

LOOP
Parameters to extract from metafor (sigma2's, s.nlevels)

@@

#### Meta-analysis of heterogeneity
```{r}
## *Perform meta-meta-analysis (3 for each of the 9 grouping terms: var.CVR, var.VR, var.RR)

@@

## Acknowledgements
tbd

## R Session Information
